\section{Limitations}

This draft has four important limitations.

First, the current RBTAD result on \liberocorelt{} is a single-seed evaluation. Beyond the Majority reports three-seed averages, so a final SOTA statement requires matching that protocol.

Second, we do not have physical robot hardware for the Real-World-LT benchmark. We therefore use \liberospatiallt{} as a simulated long-tail generalization test, but this is not a substitute for the original real-world protocol.

Third, negative instruction construction is currently task-template based. This is simple and low-cost, but it can create false negatives if two tasks share the same early action. The current mild margin and sampling ratio reduce this risk, but do not eliminate it.

Fourth, the methods improve task-conditioned action boundaries but do not solve all rare relation tasks. In particular, the wine-bottle-to-rack task remains near zero in the current \liberocorelt{} run. On \liberospatiallt{}, RSDF gives a promising seed 7 gain but only a small seed 13 gain, still leaves tasks 5 and 7 at zero, and can reduce task 6 relative to the baseline. Future work should combine task-conditioned discrimination with better state eligibility, object-relation grounding diagnostics, and multi-seed validation.

Additional Spatial-LT screens after RSDF further narrow the failure mode. Full-parameter L2-SP correction, confusion-gated rare behavior cloning, a five-step corrective pulse, and LLM-only relation-localized correction all fail to produce a consistent two-seed gain. This suggests that future methods should preserve the baseline action distribution directly while applying relation-specific corrective gradients, rather than relying on parameter anchoring, blanket rare weighting, or module freezing alone.
